A word-sense representation for documents based on information value and word sense disambiguation

### 2014-02-24 paper/refactor.tex
add wsd text with bio reference. add generated tables from code

\subsubsection{Word sense disambiguation}

\subsubsection{Top-sense definition}

@@

\appendix[All tables]

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
    \multicolumn{2}{|c|}{State and Revolution} \\ \hline
    \# & Word  \\ \hline
1 & kautsky \\ \hline
2 & marx \\ \hline
3 & revolution \\ \hline
4 & works \\ \hline
5 & engels \\ \hline
6 & society \\ \hline
7 & power \\ \hline
8 & party \\ \hline
9 & social \\ \hline
10 & experience \\ \hline
11 & bernstein \\ \hline
12 & question \\ \hline
13 & equality \\ \hline
14 & phase \\ \hline
15 & 1917 \\ \hline
16 & commune \\ \hline
17 & one \\ \hline
18 & republic \\ \hline
19 & class \\ \hline
20 & law \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
    \multicolumn{2}{|c|}{Imperialism, the Highest ...} \\ \hline
    \# & Word  \\ \hline
1 & enterprises \\ \hline
2 & bank \\ \hline
3 & imperialism \\ \hline
4 & world \\ \hline
5 & million \\ \hline
6 & colonies \\ \hline
7 & kautsky \\ \hline
8 & colonial \\ \hline
9 & production \\ \hline
10 & cent \\ \hline
11 & economic \\ \hline
12 & company \\ \hline
13 & capitalist \\ \hline
14 & monopoly \\ \hline
15 & big \\ \hline
16 & countries \\ \hline
17 & policy \\ \hline
18 & powers \\ \hline
19 & 000 \\ \hline
20 & banks \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
\multicolumn{2}{|c|}{Materialism and Empirio-criticism} \\ \hline
\# & Word  \\ \hline
1 & physics \\ \hline
2 & social \\ \hline
3 & avenarius \\ \hline
4 & space \\ \hline
5 & time \\ \hline
6 & dietzgen \\ \hline
7 & berkeley \\ \hline
8 & truth \\ \hline
9 & rey \\ \hline
10 & objective \\ \hline
11 & sensation \\ \hline
12 & experience \\ \hline
13 & pp \\ \hline
14 & engels \\ \hline
15 & motion \\ \hline
16 & marx \\ \hline
17 & book \\ \hline
18 & works \\ \hline
19 & vol \\ \hline
20 & general \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
\multicolumn{2}{|c|}{What Is To Be Done?} \\ \hline
\# & Word  \\ \hline
1 & local \\ \hline
2 & organisation \\ \hline
3 & abroad \\ \hline
4 & union \\ \hline
5 & criticism \\ \hline
6 & economic \\ \hline
7 & political \\ \hline
8 & revolutionaries \\ \hline
9 & freedom \\ \hline
10 & newspapers \\ \hline
11 & workers \\ \hline
12 & nadezhdin \\ \hline
13 & consciousness \\ \hline
14 & newspaper \\ \hline
15 & bourgeois \\ \hline
16 & masses \\ \hline
17 & martynov \\ \hline
18 & would \\ \hline
19 & movement \\ \hline
20 & german \\ \hline
  \end{tabular}
\end{center}

\begin{center}
  \begin{tabular}{ | l | l | }
    \hline
\multicolumn{2}{|c|}{The Agrarian Programme of ...} \\ \hline
\# & Word  \\ \hline
1 & rent \\ \hline
2 & dessiatins \\ \hline
3 & 000 \\ \hline
4 & land \\ \hline
5 & maslov \\ \hline
6 & peasants \\ \hline
7 & capital \\ \hline
8 & revolution \\ \hline
9 & marx \\ \hline
10 & people \\ \hline
11 & nationalisation \\ \hline
12 & plekhanov \\ \hline
13 & restoration \\ \hline
14 & right \\ \hline
15 & latifundia \\ \hline
16 & private \\ \hline
17 & farming \\ \hline
18 & free \\ \hline
19 & central \\ \hline
20 & lands \\ \hline
  \end{tabular}
\end{center}

\appendix[Entropy of a word in a text]
\label{appendix1}
Let's recall from \cite{DARWIN} and \cite{ENTROPIC} the concept of information value of a word. 
Let $T$ be a text, which has been split into $P$ pieces $T_1, T_2, \dots, T_P$, $w$ be a word in $T$,
and $f_i$ the frequency of appearance of $w$ in $T_i$.

@@

\begin{thebibliography}{10}
\bibitem{LENIN}
Marxists Internet Archive, {\em www.marxists.org} %, No estaria pudiendo definir bien esta referencia
\bibitem{DARWIN}
M.A. Montemurro and D. H. Zanette, {\em The statistics of meaning: Darwin, Gibbon and Moby Dick}, Significance, Dec. 2009, 165-169.
\bibitem{ENTROPIC}
M.A. Montemurro and D. H. Zanette, {\em Entropic analysis of the role of words in literary texts}, Adv. Complex Systems 5, September 27th 2001
\bibitem{MAS-ZANETTE}
\bibitem{WORDNET}
George A. Miller, {\em WordNet: A Lexical Database for English }, COMMUNICATIONS OF THE ACM November 1995/Vol. 38, No. 11
\bibitem{LESK}
Autor, {\em Nombre }, Publicación, No. %Buscar y poner
\end{thebibliography}
\end{article}
\end{document}

>>>>

\subsubsection{Word sense disambiguation}
\input{wsd}
\subsubsection{Top-sense definition}

@@

\appendix[All tables]
\input{top_senses_the_state_and_revolution}
\appendix[Entropy of a word in a text]
\label{appendix1}
Let's recall from \cite{DARWIN} and \cite{ENTROPIC} the concept of information value of a word.
Let $T$ be a text, which has been split into $P$ pieces $T_1, T_2, \dots, T_P$, $w$ be a word in $T$,
and $f_i$ the frequency of appearance of $w$ in $T_i$.

@@

\begin{thebibliography}{10}
\bibitem{LENIN}
Marxists Internet Archive, {\em www.marxists.org} %, No estaria pudiendo definir bien esta referencia
\bibitem{DARWIN}
M.A. Montemurro and D. H. Zanette, {\em The statistics of meaning: Darwin, Gibbon and Moby Dick}, Significance, Dec. 2009, 165-169.
\bibitem{ENTROPIC}
M.A. Montemurro and D. H. Zanette, {\em Entropic analysis of the role of words in literary texts}, Adv. Complex Systems 5, September 27th 2001
\bibitem{MAS-ZANETTE}
\bibitem{WORDNET}
George A. Miller, {\em WordNet: A Lexical Database for English }, COMMUNICATIONS OF THE ACM November 1995/Vol. 38, No. 11
\bibitem{LESK}
M. Lesk. Automatic sense disambiguation using machine readable dictionaries: How to tell a pine cone from a ice cream cone. In Proceedings of SIGDOC ’86, 1986.
\end{thebibliography}
\end{article}
\end{document}

### 2014-02-24 paper/wsd.tex
add wsd text with bio reference. add generated tables from code

Our approach for WSD was a classical Lesk implementation that uses wordnet as a definition source with some tweaks.
Basically Lesk is based on the idea that words that co–occur in a sentence are being used to refer to the same topic.
The Lesk alghorithm{BIO} takes the word definition or gloss that we want to disambiguate and it compares it gloss against the glosses of the other words in the context.
Then a word is assigned that sense whose gloss shares the largest number of words in common with the glosses of the other words.
This algorhitm suffers the problem that relies on the word definition and definitions usually uses as less words as possible.
Not only that sometimes two different words, but related ones, uses no words in common in their definitions.
As we use the nltk wordnet we immediately notice an improvement using the rich set of relationships that Wordnet gives to us.
We added to Lesk word stems, hypernyms, hyponyms.
For the problem of short word description we modified the algorthim to use the definition of the word to disambiguate and to compare it against the context words directly.
Finally our algorithm chooses the best synset from wordnet based on the context words.

BIO
M. Lesk. Automatic sense disambiguation using machine readable dictionaries: How to tell a pine cone from a ice cream cone. In Proceedings of SIGDOC ’86, 1986.

>>>>

Our approach for WSD was a classical Lesk implementation that uses wordnet as a definition source with some tweaks.
Basically Lesk is based on the idea that words that co–occur in a sentence are being used to refer to the same topic.
The Lesk alghorithm \cite{LESK} takes the word definition or gloss that we want to disambiguate and it compares it gloss against the glosses of the other words in the context.
Then a word is assigned that sense whose gloss shares the largest number of words in common with the glosses of the other words.
This algorhitm suffers the problem that relies on the word definition and definitions usually uses as less words as possible.
Not only that sometimes two different words, but related ones, uses no words in common in their definitions.
As we use the nltk wordnet we immediately notice an improvement using the rich set of relationships that Wordnet gives to us.
We added to Lesk word stems, hypernyms, hyponyms.
For the problem of short word description we modified the algorthim to use the definition of the word to disambiguate and to compare it against the context words directly.
Finally our algorithm chooses the best synset from wordnet based on the context words.

BIO
